Streaking and Clutching in Round-Based First Person Shooter Esports

### 2022-07-17 presentation.qmd
reformat bullets
make a good looking pdf with renderthis

---
title: "Streaking and Clutching in Round-Based First Person Shooter (FPS) Esports"
author: "Tony ElHabr"
format:
  revealjs:
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

```{r}
#| label: setup-eval
library(tidyverse)
library(extrafont)
library(ggtext)
library(ggsci)
blackish_background <- '#191919' ## to match slide background
gray_points <- '#4d4d4d'
gray_text <- '#999999'

@@

common_cod_labs <- function(...) {
  list(
    ...,
    labs(
      caption = 'CDL SnD Major and Weekly matches, 2020 - present',
      tag = '**Viz**: @TonyElHabr | **Data**: @IOUTurtle'
    )
  )
}

@@

.dpi <- 300
save_heatmap <- function(
    p, 
    filename = paste0(str_remove(deparse(substitute(p)), '^p_'), '.png'), 
    width = 7.5, 
    height = 5,
    units = 'px',
    dpi = .dpi,
    ...
) {
  ggsave(
    plot = p,
    filename = filename,
    width = width * dpi,
    height = height * dpi,
    units = units,
    dpi = dpi,
    ...
  )
}

@@

p_o_round_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  heatmap_text_layers(.f = o_round_win_text_layer) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_o_round_win_prop
save_heatmap(p_o_round_win_prop)
```

@@

max_diff_win_round_prop <- max(abs(o_win_prop$diff_win_round_prop))
p_o_round_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'diff_win_round_prop') +
  heatmap_text_layers(.f = o_round_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_round_prop, 
      +max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_o_round_win_prop_over_expected
save_heatmap(p_o_round_win_prop_over_expected)
```

@@

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('deep-orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_heatmap(p_o_series_win_prop)
```

@@

max_diff_win_series_prop <- max(abs(o_win_prop$diff_win_series_prop))
p_o_series_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  pre_cumu_heatmap_tile(.fill = 'diff_win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_series_prop, 
      +max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_o_series_win_prop_over_expected
save_heatmap(p_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: p_o_win_r1_prop
calculate_o_round_win_prop <- function(df) {
  df |> 
    count(year, game, map, is_offense, win_round) |> 
    group_by(map, is_offense) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |>
    filter(win_round, is_offense) |>
    select(
      year,
      game,
      map,
      rounds_won = n,
      total_rounds = total,
      win_round_prop = prop
    ) |>
    arrange(desc(win_round_prop))
}

devtools::source_url('https://raw.githubusercontent.com/dgrtwo/splittestr/master/R/vectorized-prop-test.R')
summarize_o_round_win_prop <- function(.round) {

@@

p_o_win_r1_vs_post_round_prop <- res_r1$plot +
  geom_curve(
    data = tibble(),
    aes(x = 1.2, y = 0.65, xend = 1.1, yend = 0.63),
    curvature = 0.2, 
    color = o_win_rx_prop_pal[["-1"]], 
    arrow = arw
  ) +
  geom_curve(
    data = tibble(),
    aes(x = 1.24, y = 0.27, xend = 1.1, yend = 0.27),
    curvature = 0-.2, 
    color = o_win_rx_prop_pal[["1"]], 
    arrow = arw
  ) +
  geom_text(
    data = tibble(
      x = c(1.2, 1.24),
      y = c(0.65, 0.27),
      hjust = c(0, 0),
      vjust = c(0.5, 0.5),
      lab = c(
        sprintf(
          'Largest negative percent difference in round 1\nvs. all other round win percentages (%+.0f%%).', 
          100 * res_r1$largest_neg_diff_prop
        ),
        sprintf(
          'Largest positive percent difference (%+.0f%%).\nOnly significant difference (p value < 0.01).',
          100 * res_r1$largest_pos_diff_prop
        )
      )
    ),
    color = 'white',
    size = 9 / .pt,
    aes(x = x, y = y, hjust = hjust, vjust = vjust, label = lab)
  )
p_o_win_r1_vs_post_round_prop

save_slope_chart <- function(
    p, 
    filename = paste0(str_remove(deparse(substitute(p)), '^p_'), '.png'), 
    width = 7.5, 
    height = 5,
    units = 'px',
    dpi = .dpi,
    ...
) {

  ggsave(
    plot = p,
    filename = filename,
    width = 7.5 * .dpi,
    height = 5 * .dpi,
    units = units,
    dpi = dpi,
    ...
  )
}

save_slope_chart(p_o_win_r1_vs_post_round_prop)
```

@@

```{r}
res_r2 <- plot_round_vs_post_round_slope_chart(2)
res_r2$plot
save_slope_chart(
  res_r2$plot,
  filename = 'o_win_r2_vs_post_round_prop.png'
)
```

@@

## How do end states compare to expectation?

```{r}
#| label: series_outcomes
prior_clinching_round_win_prop <- rounds |> 
  filter(round == (n_rounds - 1L)) |> 
  count(n_rounds, win_series, won_prior_round = win_round) |> 
  mutate(
    across(win_series, ~ifelse(.x, 'win', 'loss'))
  ) |> 
  pivot_wider(
    names_from = win_series,
    values_from = n,
    values_fill = 0L
  ) |> 
  filter(won_prior_round) |> 
  select(-won_prior_round) |> 
  mutate(
    win_series_prop = win / (win + loss),
    e_win_series_prop = (6L - 1L) / (n_rounds - 1L)
  )

@@

series_streaks <- full_join(
  actual_series_streaks |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  e_series_streaks |> 
    rename_with(~sprintf('%s_expected', .x), c(n, prop))
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected
  ) |> 
  arrange(desc(abs(prop_diff)))
series_streaks |> filter(is.na(n_actual))

## 6-0, 6-1, and some combos of 6-5 occur more often than expected
series_streaks |> 
  filter(!is.na(n_actual)) |> 
  mutate(total_actual = sum(n_actual)) |> 
  mutate(
    p = vectorized_prop_test(n_actual, total_actual, n_expected, !!n_sims)
  ) |> 
  select(-total_actual) |> 
  arrange(p$p.value)

## 6-0 and 6-1 occur more often than expected
series_outcomes <- full_join(
  actual_series_streaks |> 
    group_by(record) |> 
    summarize(
      across(n, sum)
    ) |> 
    mutate(prop = n / sum(n)) |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  e_series_streaks |> 
    group_by(record) |> 
    summarize(
      across(n, sum)
    ) |> 
    mutate(prop = n / sum(n)) |> 
    rename_with(~sprintf('%s_expected', .x), c(n, prop))
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected
  ) |> 
  arrange(desc(abs(prop_diff))) |> 
  mutate(total_actual = sum(n_actual)) |> 
  mutate(
    p = vectorized_prop_test(n_actual, total_actual, n_expected, !!n_sims)
  ) |> 
  select(-total_actual) |> 
  arrange(p$p.value)

series_outcomes |> 
  select(record, actual = prop_actual, expected = prop_expected, diff = prop_diff) |> 
  pivot_longer(
    -record,
    names_to = 'prop',
    values_to = 'value'
  ) |> 
  mutate(
    across(
      prop,
      factor, 
      levels = c('actual', 'expected', 'diff')
    )
  ) |> 
  ggplot() +
  aes(x = record, y = prop) +
  geom_tile(
    show.legend = FALSE,
    aes(fill = value)
  ) +
  scale_x_discrete(
    position = 'top'
  )

```

::: notes
-   There should be evidence of 6-0's and 6-5's being more frequent than expected given 50-50% chances.
:::

@@

```{r}
summarize_b2b_round_wins <- function(df, ...) {
  df |> 
    arrange(year, sheet, series, team, round) |> 
    group_by(year, sheet, series, team) |> 
    mutate(
      b2b_round_wins = win_round & lag(win_round)
    ) |> 
    ungroup() |> 
    filter(round > 1L) |> 
    count(..., round, b2b_round_wins) |> 
    group_by(..., round) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |> 
    filter(b2b_round_wins) |> 
    select(-b2b_round_wins)
}

## now accounting for number of rounds that end up being played...
## seems like there is interesting behavior in the last round, regardless `n_rounds`
b2b_round_win_props <- rounds |> 
  filter(n_rounds > 6L) |> 
  summarize_b2b_round_wins(n_rounds) |> 
  group_by(n_rounds) |> 
  mutate(
    z = (n - mean(n)) / sd(n),
    rnk = row_number(desc(n))
  ) |> 
  ungroup()

b2b_round_win_props_tile_layer <- function(..., .round, .palette, .is_new = TRUE) {

@@

  list(
    ...,
    init,
    geom_tile(
      data = b2b_round_win_props |> filter(n_rounds == .round),
      color = blackish_background,
      height = 0.8,
      show.legend = FALSE,
      aes(
        fill = z,
        x = round + 0.5, 
        y = n_rounds + 0.5
      )
    ),
    ggsci::scale_fill_material(palette = .palette) 
  )
}

b2b_round_win_prop_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `>`
    .color <- blackish_background
  } else {
    op <- `==`
    .color <- 'white'
  }
  list(
    ...,
    geom_text(
      data = b2b_round_win_props |> filter(`op`(rnk, 1)),
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        label = sprintf('%s\n(%s)', scales::percent(prop, accuracy = 1), scales::comma(n)),
        x = round + 0.5, 
        y = n_rounds + 0.5
      )
    )
  )
}

@@

b2b_round_win_props_arws <- tibble(
  x = c(9.3, 9.4, 9.75, 10),
  y = c(8.25, 8.5, 8.75, 8.75),
  xend = c(8.1, 9.1, 9.5, 10.5),
  yend = c(7.5, 8.5, 9, 10),
  curvature = c(-0.2, -0.1, 0.1, 0.2)
) |> 
  mutate(
    rn = row_number()
  )

p_b2b_round_win_props <- b2b_round_win_props |> 
  ggplot() +
  b2b_round_win_props_tile_layer(.round = 7, .palette = 'pink', .is_new = FALSE) +
  b2b_round_win_props_tile_layer(.round = 8, .palette = 'light-blue') +
  b2b_round_win_props_tile_layer(.round = 9, .palette = 'light-green') +
  b2b_round_win_props_tile_layer(.round = 10, .palette = 'amber') +
  b2b_round_win_props_tile_layer(.round = 11, .palette = 'deep-orange') +
  b2b_round_win_prop_text_layers() +
  scale_x_continuous(
    labels = 2:11,
    breaks = seq(2.5, 11.5, by = 1),
    expand = c(0, 0.5)
  ) +
  scale_y_continuous(
    labels = 7:11,
    breaks = seq(7.5, 11.5, by = 1),
    expand = c(0, 0)
  ) +
  theme(
    panel.grid.major = element_blank(),
    axis.title = element_text(hjust = 0.5),
    plot.title = ggtext::element_markdown(hjust = 0.5),
    axis.text = element_text(size = 16, face = 'bold')
  ) +
  common_cod_labs() +
  labs(
    title = 'Back-to-back round win %',
    x = 'Round',
    y = '# of Rounds in Complete Series'
  ) + 
  geom_text(
    data = tibble(
      label = 'Back-to-back wins occur most\nfrequently in the last round.',
      x = 9.5,
      y = 8.5
    ),
    size = 9 / .pt,
    fontface = 'bold',
    color = 'white',
    hjust = 0,
    vjust = 0.5,
    aes(x = x, y = y, label = label)
  ) +
  map(
    b2b_round_win_props_arws |> group_split(rn), 
    function(.x) {
      geom_curve(
        data = data.frame(),
        color = 'white',
        arrow = arw,
        aes(x = .x$x, y = .x$y, xend = .x$xend, yend = .x$yend),
        curvature = .x$curvature
      )
    }
  )
save_heatmap(p_b2b_round_win_props)
# ## now accounting for number of rounds that end up being played AND series winners...
# rounds |> 
#   filter(win_series, n_rounds > 6L) |> 
#   summarize_b2b_round_wins(n_rounds)
# 
# ## and the losers...
# rounds |> 
#   filter(!win_series) |> 
#   summarize_b2b_round_wins(n_rounds)
```

![](b2b_round_win_props.png){fig-align="center"}

@@

>>>>

---
title: "Streaking and Clutching in Round-Based First Person Shooter Esports"
author: "Tony ElHabr"
format:
  revealjs:
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

@@

```{r}
#| label: setup-eval
library(tidyverse)
library(extrafont)
library(ggtext)
library(ggsci)
library(ggnewscale)
blackish_background <- '#191919' ## to match slide background
gray_points <- '#4d4d4d'
gray_text <- '#999999'

@@

common_cod_labs <- function(...) {
  list(
    ...,
    labs(
      caption = 'CDL SnD Major and Weekly matches, 2020 - present'
    )
  )
}

@@

.dpi <- 300
save_for_slide <- function(
    p, 
    filename = paste0(str_remove(deparse(substitute(p)), '^p_'), '.png'), 
    width = 7.5, 
    height = width / 1.5,
    units = 'px',
    dpi = .dpi,
    ...
) {
  ggsave(
    plot = p,
    filename = filename,
    width = width * dpi,
    height = height * dpi,
    units = units,
    dpi = dpi,
    ...
  )
}

@@

p_o_round_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  heatmap_text_layers(.f = o_round_win_text_layer) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_o_round_win_prop
save_for_slide(p_o_round_win_prop)
```

@@

max_diff_win_round_prop <- max(abs(o_win_prop$diff_win_round_prop))
p_o_round_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_round_prop') +
  heatmap_text_layers(.f = o_round_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_round_prop, 
      +max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_o_round_win_prop_over_expected
save_for_slide(p_o_round_win_prop_over_expected)
```

@@

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('deep-orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_for_slide(p_o_series_win_prop)
```

@@

max_diff_win_series_prop <- max(abs(o_win_prop$diff_win_series_prop))
p_o_series_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_cod_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_series_prop') +
  heatmap_text_layers(.f = add_o_series_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_series_prop, 
      +max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_o_series_win_prop_over_expected
save_for_slide(p_o_series_win_prop_over_expected)
```

@@

```{r}
#| label: o_win_r1_vs_post_round_pro
calculate_o_round_win_prop <- function(df) {
  df |> 
    count(year, game, map, is_offense, win_round) |> 
    group_by(map, is_offense) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |>
    filter(win_round, is_offense) |>
    select(
      year,
      game,
      map,
      rounds_won = n,
      total_rounds = total,
      win_round_prop = prop
    ) |>
    arrange(desc(win_round_prop))
}

# https://raw.githubusercontent.com/dgrtwo/splittestr/master/R/vectorized-prop-test.R
vectorized_prop_test_approx <- function(a, b, c, d) {
  n1 <- a + b
  n2 <- c + d
  n <- n1 + n2
  p <- (a + c) / n
  E <- cbind(p * n1, (1 - p) * n1, p * n2, (1 - p) * n2)

  x <- cbind(a, b, c, d)

  DELTA <- a / n1 - c / n2
  YATES <- pmin(.5, abs(DELTA) / sum(1 / n1 + 1 / n2))

  STATISTIC <- rowSums((abs(x - E) - YATES)^2 / E)
  PVAL <- pchisq(STATISTIC, 1, lower.tail = FALSE)
  PVAL
}

vectorized_prop_test_exact <- function(a, b, c, d) {
  sapply(seq_along(a), function(i) {
    fisher.test(cbind(c(a[i], c[i]), c(b[i], d[i])))$p.value
  })
}

vectorized_prop_test <- function(x1, n1, x2, n2, conf.level = .95) {
  a <- x1
  b <- n1 - x1
  c <- x2
  d <- n2 - x2

  # if any values are < 20, use Fisher's exact test
  exact <- (a < 20 | b < 20 | c < 20 | d < 20)

  pvalue <- rep(NA, length(a))

  if (any(exact)) {
    pvalue[exact] <- vectorized_prop_test_exact(a[exact], b[exact], c[exact], d[exact])
  }
  if (any(!exact)) {
    pvalue[!exact] <- vectorized_prop_test_approx(a[!exact], b[!exact], c[!exact], d[!exact])
  }

  mu1 <- a / (a + b)
  mu2 <- c / (c + d)

  ## confidence interval
  alpha2 <- (1 - conf.level) / 2
  DELTA <- mu2 - mu1
  WIDTH <- qnorm(alpha2)
  alpha <- (a + .5) / (a + b + 1)
  beta <- (c + .5) / (c + d + 1)

  n <- n1 + n2
  YATES <- pmin(.5, abs(DELTA) / sum(1 / n1 + 1 / n2))

  z <- qnorm((1 + conf.level) / 2)
  WIDTH <- z * sqrt(mu1 * (1 - mu1) / n1 + mu2 * (1 - mu2) / n2)

  tibble(
    estimate = DELTA,
    conf.low = pmax(DELTA - WIDTH, -1),
    conf.high = pmin(DELTA + WIDTH, 1),
    p.value = pvalue
  )
}

summarize_o_round_win_prop <- function(.round) {

@@

p_o_win_r1_vs_post_round_prop <- res_r1$plot +
  geom_curve(
    data = tibble(),
    aes(x = 1.2, y = 0.65, xend = 1.1, yend = 0.63),
    curvature = 0.2, 
    color = o_win_rx_prop_pal[["-1"]], 
    arrow = arw
  ) +
  geom_curve(
    data = tibble(),
    aes(x = 1.24, y = 0.27, xend = 1.1, yend = 0.27),
    curvature = 0-.2, 
    color = o_win_rx_prop_pal[["1"]], 
    arrow = arw
  ) +
  geom_text(
    data = tibble(
      x = c(1.2, 1.24),
      y = c(0.65, 0.27),
      hjust = c(0, 0),
      vjust = c(0.5, 0.5),
      lab = c(
        sprintf(
          'Largest negative percent difference in round 1\nvs. all other round win percentages (%+.0f%%).', 
          100 * res_r1$largest_neg_diff_prop
        ),
        sprintf(
          'Largest positive percent difference (%+.0f%%).\nOnly significant difference (p value < 0.01).',
          100 * res_r1$largest_pos_diff_prop
        )
      )
    ),
    color = 'white',
    size = 9 / .pt,
    aes(x = x, y = y, hjust = hjust, vjust = vjust, label = lab)
  )
p_o_win_r1_vs_post_round_prop
save_for_slide(p_o_win_r1_vs_post_round_prop)
```

@@

```{r}
#| label: o_win_r2_vs_post_round_pro
res_r2 <- plot_round_vs_post_round_slope_chart(2)
res_r2$plot
save_for_slide(
  res_r2$plot,
  filename = 'o_win_r2_vs_post_round_prop.png'
)
```

@@

```{r}
#| label: p_series_outcome_prop_over_expected
prior_clinching_round_win_prop <- rounds |> 
  filter(round == (n_rounds - 1L)) |> 
  count(n_rounds, win_series, won_prior_round = win_round) |> 
  mutate(
    across(win_series, ~ifelse(.x, 'win', 'loss'))
  ) |> 
  pivot_wider(
    names_from = win_series,
    values_from = n,
    values_fill = 0L
  ) |> 
  filter(won_prior_round) |> 
  select(-won_prior_round) |> 
  mutate(
    win_series_prop = win / (win + loss),
    e_win_series_prop = (6L - 1L) / (n_rounds - 1L)
  )

@@

series_streaks <- full_join(
  actual_series_streaks |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  e_series_streaks |> 
    rename_with(~sprintf('%s_expected', .x), c(n, prop))
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected
  ) |> 
  arrange(desc(abs(prop_diff)))
# series_streaks |> filter(is.na(n_actual))

## 6-0, 6-1, and some combos of 6-5 occur more often than expected
series_streak_props <- series_streaks |> 
  filter(!is.na(n_actual)) |> 
  mutate(total_actual = sum(n_actual)) |> 
  mutate(
    p = vectorized_prop_test(n_actual, total_actual, n_expected, !!n_sims)
  ) |> 
  select(-total_actual) |> 
  arrange(p$p.value)

## 6-0 and 6-1 occur more often than expected
series_outcome_prop <- full_join(
  actual_series_streaks |> 
    group_by(record) |> 
    summarize(
      across(n, sum)
    ) |> 
    mutate(prop = n / sum(n)) |> 
    rename_with(~sprintf('%s_actual', .x), c(n, prop)),
  e_series_streaks |> 
    group_by(record) |> 
    summarize(
      across(n, sum)
    ) |> 
    mutate(prop = n / sum(n)) |> 
    rename_with(~sprintf('%s_expected', .x), c(n, prop))
) |> 
  mutate(
    prop_diff = prop_actual - prop_expected,
    total_actual = sum(n_actual),
    n_expected_downsampled = round(prop_expected * total_actual),
    p = vectorized_prop_test(n_actual, total_actual, n_expected_downsampled, total_actual)
  ) |> 
  select(-total_actual) |> 
  arrange(p$p.value)

long_series_outcome_prop <- series_outcome_prop |>
  unnest(p) |>
  select(
    record,
    n = n_actual,
    `3` = prop_actual,
    `2` = prop_expected,
    `1` = prop_diff
  ) |> 
  pivot_longer(
    matches('[123]'),
    names_to = 'outcome_idx',
    values_to = 'value'
  ) |> 
  mutate(across(outcome_idx, as.integer))

max_diff_series_outcome_prop <- long_series_outcome_prop |> 
  filter(outcome_idx == 1) |> 
  pull(value) |> 
  abs() |> 
  max()

p_series_outcome_prop_over_expected <- long_series_outcome_prop |> 
  ggplot() +
  aes(x = record, y = outcome_idx) +
  geom_tile(
    data = long_series_outcome_prop |> filter(outcome_idx > 1),
    show.legend = FALSE,
    # height = 0.8,
    aes(fill = value)
  ) +
  ggsci::scale_fill_material(
    palette = 'orange'
  ) +
  ggnewscale::new_scale_fill() +
  geom_tile(
    data = long_series_outcome_prop |> filter(outcome_idx == 1),
    show.legend = FALSE,
    height = 0.8,
    aes(fill = value)
  ) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_series_outcome_prop, 
      +max_diff_series_outcome_prop
    )
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 3),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%.1f%%\n(%s)', 100 * value, scales::comma(n)))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 2),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%.1f%%', 100 * value))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 1) |> filter(abs(value) != max(abs(value))),
    color = blackish_background,
    size = 11 / .pt,
    aes(label = sprintf('%+.1f%%', 100 * value))
  ) +
  geom_text(
    data = long_series_outcome_prop |> filter(outcome_idx == 1) |> filter(abs(value) == max(abs(value))),
    color = 'white',
    size = 11 / .pt,

    aes(label = sprintf('%+.1f%%', 100 * value))
  ) +
  scale_x_discrete(
    position = 'top',
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    labels = c('Diff.', 'Expected', 'Actual'),
    breaks = 1:3,
    expand = c(0, 0)
  ) +
  theme(
    panel.grid.major = element_blank(),
    axis.title = element_text(hjust = 0.5),
    plot.title = ggtext::element_markdown(hjust = 0.5),
    axis.text = element_text(size = 16, face = 'bold')
  ) +
  labs(
    title = 'Series outcome %',
    x = NULL,
    y = NULL
  ) +
  common_cod_labs() 
p_series_outcome_prop_over_expected
save_for_slide(p_series_outcome_prop_over_expected)
```

![](series_outcome_prop_over_expected.png){fig-align="center"}

::: notes
-   There should be evidence of 6-0's and 6-1's are more frequent than expected given 50-50% chances.
-   Notably 6-4's occur less frequently than expected
-   p-values are not significant for any
:::

@@

```{r}
#| label: p_b2b_round_win_prop
summarize_b2b_round_wins <- function(df, ...) {
  df |> 
    arrange(year, sheet, series, team, round) |> 
    group_by(year, sheet, series, team) |> 
    mutate(
      b2b_round_wins = win_round & lag(win_round)
    ) |> 
    ungroup() |> 
    filter(round > 1L) |> 
    count(..., round, b2b_round_wins) |> 
    group_by(..., round) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |> 
    filter(b2b_round_wins) |> 
    select(-b2b_round_wins)
}

## now accounting for number of rounds that end up being played...
## seems like there is interesting behavior in the last round, regardless `n_rounds`
b2b_round_win_prop <- rounds |> 
  filter(n_rounds > 6L) |> 
  summarize_b2b_round_wins(n_rounds) |> 
  group_by(n_rounds) |> 
  mutate(
    z = (n - mean(n)) / sd(n),
    rnk = row_number(desc(n))
  ) |> 
  ungroup()

b2b_round_win_prop_tile_layer <- function(..., .round, .palette, .is_new = TRUE) {

@@

  list(
    ...,
    init,
    geom_tile(
      data = b2b_round_win_prop |> filter(n_rounds == .round),
      color = blackish_background,
      height = 0.8,
      show.legend = FALSE,
      aes(
        fill = z,
        x = round + 0.5, 
        y = n_rounds + 0.5
      )
    ),
    ggsci::scale_fill_material(palette = .palette) 
  )
}

b2b_round_win_prop_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `>`
    .color <- blackish_background
  } else {
    op <- `==`
    .color <- 'white'
  }
  list(
    ...,
    geom_text(
      data = b2b_round_win_prop |> filter(`op`(rnk, 1)),
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        label = sprintf('%s\n(%s)', scales::percent(prop, accuracy = 1), scales::comma(n)),
        x = round + 0.5, 
        y = n_rounds + 0.5
      )
    )
  )
}

@@

b2b_round_win_prop_arws <- tibble(
  x = c(9.3, 9.4, 9.75, 10),
  y = c(8.25, 8.5, 8.75, 8.75),
  xend = c(8.1, 9.1, 9.5, 10.5),
  yend = c(7.5, 8.5, 9, 10),
  curvature = c(-0.2, -0.1, 0.1, 0.2)
) |> 
  mutate(
    rn = row_number()
  )

p_b2b_round_win_prop <- b2b_round_win_prop |> 
  ggplot() +
  b2b_round_win_prop_tile_layer(.round = 7, .palette = 'pink', .is_new = FALSE) +
  b2b_round_win_prop_tile_layer(.round = 8, .palette = 'light-blue') +
  b2b_round_win_prop_tile_layer(.round = 9, .palette = 'light-green') +
  b2b_round_win_prop_tile_layer(.round = 10, .palette = 'amber') +
  b2b_round_win_prop_tile_layer(.round = 11, .palette = 'deep-orange') +
  b2b_round_win_prop_text_layers() +
  scale_x_continuous(
    labels = 2:11,
    breaks = seq(2.5, 11.5, by = 1),
    expand = c(0, 0.5)
  ) +
  scale_y_continuous(
    labels = 7:11,
    breaks = seq(7.5, 11.5, by = 1),
    expand = c(0, 0)
  ) +
  theme(
    panel.grid.major = element_blank(),
    axis.title = element_text(hjust = 0.5),
    plot.title = ggtext::element_markdown(hjust = 0.5),
    axis.text = element_text(size = 16, face = 'bold')
  ) +
  common_cod_labs() +
  labs(
    title = 'Back-to-back round win %',
    x = 'Round',
    y = '# of Rounds in Complete Series'
  ) + 
  geom_text(
    data = tibble(
      label = 'Back-to-back wins occur most\nfrequently in the last round.',
      x = 9.5,
      y = 8.5
    ),
    size = 9 / .pt,
    fontface = 'bold',
    color = 'white',
    hjust = 0,
    vjust = 0.5,
    aes(x = x, y = y, label = label)
  ) +
  map(
    b2b_round_win_prop_arws |> group_split(rn), 
    function(.x) {
      geom_curve(
        data = data.frame(),
        color = 'white',
        arrow = arw,
        aes(x = .x$x, y = .x$y, xend = .x$xend, yend = .x$yend),
        curvature = .x$curvature
      )
    }
  )
save_for_slide(p_b2b_round_win_prop)
# ## now accounting for number of rounds that end up being played AND series winners...
# rounds |> 
#   filter(win_series, n_rounds > 6L) |> 
#   summarize_b2b_round_wins(n_rounds)
# 
# ## and the losers...
# rounds |> 
#   filter(!win_series) |> 
#   summarize_b2b_round_wins(n_rounds)
```

![](b2b_round_win_prop.png){fig-align="center"}

@@

```{r}
#| label: plant_win_prop_proj
map_plant_win_prop <- rounds |> 
  filter(!is.na(plant)) |> 
  filter(is_offense) |> 
  group_by(year, game, map, plant) |> 
  summarize(
    n = n(),
    wins = sum(win_round)
  ) |> 
  ungroup() |> 
  mutate(
    win_prop = wins / n
  ) |> 
  group_by(year, game, map) |> 
  mutate(
    total = sum(n),
    rnk = row_number(-n),
    prop = n / total
  ) |> 
  ungroup()

plant_win_prop <- map_plant_win_prop |>
  mutate(site = ifelse(rnk == 1, 'pri', 'sec')) |> 
  group_by(site) |> 
  summarize(
    across(c(n, wins, total), sum)
  ) |> 
  ungroup() |> 
  mutate(
    win_prop = wins / n,
    prop = n / total
  )

plant_win_prop_proj <- plant_win_prop |> 
  left_join(
    rounds |> 
      filter(is_offense) |> 
      filter(is.na(plant)) |> 
      summarize(
        n_no_plant = n(),
        wins_no_plant = sum(win_round)
      ),
    by = character()
  ) |> 
  mutate(
    n_proj = round(n + prop * n_no_plant),
    wins_proj = round(wins + prop * wins_no_plant),
    win_prop_proj = wins_proj / n_proj
  )

plant_site_prop <- plant_win_prop_proj |> 
  select(site, prop) |> 
  deframe() |> 
  scales::percent(accuracy = 0.1)

plant_site_win_prop <- plant_win_prop_proj |> 
  select(site, win_prop_proj) |> 
  deframe() |> 
  scales::percent(accuracy = 0.1)
```

```{r}
#| label: plant_win_prop_proj-eval
#| eval: true
plant_site_prop <- c(
  'pri' = '60.9%',
  'sec' = '39.1%'
)
plant_site_win_prop <- c(
  'pri' = '47.4%',
  'sec' = '48.5%'
)
```

::: columns
::: {.column width="15%"}
:::

::: {.column width="30%"}
| Plant side |                           \% |
|------------|-----------------------------:|
| Primary    | `r plant_site_prop[['pri']]` |
| Secondary  | `r plant_site_prop[['sec']]` |
:::

::: {.column width="10%"}
:::

::: {.column width="30%"}
| Plant side |                            Win % |
|------------|---------------------------------:|
| Primary    | `r plant_site_win_prop[['pri']]` |
| Secondary  |     `r plant_site_prop[['sec']]` |
:::

::: {.column width="15%"}
:::
:::
